---
title: "ESM 206 lab 5"
author: "Amelia Ritger"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to parse date with 'lubridate'
- Get counts fo observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_tile()

```{r, include=FALSE}
library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"),
                              na="-99999") %>%
  janitor::clean_names()
```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format
```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)
```

We're only going to consider 'site' as our variable of interest
```{r}
ggplot(lobster_tidy, aes(x=site, y=size_mm)) +
  geom_jitter(aes(color=site),
              alpha=0.5,
              width=0.2)

# Histogram
ggplot(lobster_tidy, aes(x=size_mm)) +
  geom_histogram(aes(fill=site)) +
  facet_wrap(~site, scales="free") #separate out histograms by site, and allow scale to fit the data rather than same scale for each site

ggplot(lobster_tidy, aes(sample=size_mm)) +
  geom_qq() +
  geom_qq_line() +
  facet_wrap(~site)

```

#### Convert the "date" column to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month & year
```{r}
lobster_date <- lobster_tidy %>%
  mutate(date_new=lubridate::mdy(date))
```

Now let's parse year and month using lubridate::month() and lubridate::year()
```{r}
lobster_parse_date <- lobster_date %>%
  mutate(obs_month=lubridate::month(date_new, label=TRUE),
         obs_year=lubridate::year(date_new)) #label converts month number to month word
```
Fun fact: (lubridate) knows to make dates ordered factors!

Now let's find counts of observed lobsters basde on different hierarchical groupings. First, let's count lobsters by year and month:

```{r}
lobster_parse_date %>%
  dplyr::count(obs_year,obs_month)

lobster_parse_date %>%
  dplyr::count(obs_year)

lobster_parse_date %>%
  dplyr::count(site)
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()
```{r}
lobster_summary <- lobster_parse_date %>%
  group_by(site) %>%
  summarize(mean_size=mean(size_mm, na.rm=TRUE),
            sd_size = sd(size_mm, na.rm = TRUE), 
            lobster_number=n())
```

#### Find confidence intervals

Use t.test() function to find CI (for one sample) and perofmr t-tests to compare means fo two samples
```{r}
ivee_lob <- lobster_tidy %>%
  filter(site=="IVEE") %>%
  pull(size_mm) #converts df into a vector 

t.test(ivee_lob)
  
```

#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs?

We've done our necessary exploratory analyses to determine that a two-sample t-test for means comparison is appropriate
```{r}
napl_sample <- lobster_tidy %>%
  filter(site=="NAPL") %>%
  pull(size_mm)

mohk_sample <- lobster_tidy %>%
  filter(site=="MOHK") %>%
  pull(size_mm)

t.test(mohk_sample, napl_sample)

```
There is a significant difference in lobster lengths between Naples and Mohawk reef (t(1851)=-19.85, p < 0.001, alpha = 0.05)

```{r}
lobster_mn <- lobster_tidy %>%
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm~site, data=lobster_mn)

mn_ttest2$statistic
mn_ttest2$parameter
mn_ttest2$p.value

```
Here is the t-statistic: `r mn_ttest2$statistic` and p-value `r mn_ttest2$p.value`

### Now, onto the heatmap!
```{r}
lobster_ys <- lobster_parse_date %>%
  count(obs_year,site)

ggplot(data=lobster_ys, aes(x=obs_year, y=site)) +
  geom_tile(aes(fill=n))
```

